Automated Topic Naming

### 2011-08-11 ese/ese.tex

\begin{abstract}
% %\todo[inline]{Topics and LDA sooner, Topics/concept location is a known
Researchers have employed a variety of techniques to extract underlying
topics that relate to software development artifacts.  Typically, these
techniques use semi-unsupervised machine-learning algorithms to suggest
candidate word-lists.  However, word-lists are
difficult to 
interpret in the absence of meaningful summary labels.  Current topic modeling 
techniques assume manual labelling and do not use domain-specific knowledge
to improve, contextualize, or describe results for the developers.  
We propose a solution: \emph{automated labelled topic extraction}.
  Topics are extracted 
using  Latent Dirichlet Allocation (LDA) from commit-log comments recovered
from source control systems such as CVS and BitKeeper.  
These topics are given labels from a generalizable
cross-project taxonomy, consisting of non-functional
requirements. Our approach was evaluated with experiments and case studies
on two large-scale RDBMS projects: MySQL and MaxDB.  The case studies
show that labelled topic extraction can produce appropriate,
context-sensitive labels relevant to these projects, which provides fresh
insight into their evolving software development activities.
%160 words
\keywords{ Software maintenance \and Repository mining \and Latent dirichlet allocation \and Topic models}

@@

To evaluate our approach, we sought candidate systems that were mature projects and had openly accessible source control repositories. 
We selected systems from the same application domain, to control for differences in functional, rather than non-functional, requirements. 
We used three different open-source, partially-commercial database systems:
\begin{itemize}
\item  MySQL 3.23. Started in 1994 and MySQL 3.23 was released in early 2001. MySQL contains $320,000$ lines of C and C++ source code\footnote{generated using David A. Wheeler's \emph{SLOCCount}, {http://dwheeler.com/sloccount}.}.
\item MaxDB 7.500. Started in the late 1970s as a research project, and was later acquired by SAP. As of version 7.500, released April 2007, the project has over $940,000$ lines of C source code. 
\item PostgreSQL 7.3. Started in the 1980s as a Berkeley research project. PostgreSQL 7.3 contains $306,000$ lines of C code.
\end{itemize}

%Choosing an older version of these projects allowed us to focus on projects which have moved further into the maintenance phase of the software life-cycle.
We explicitly chose older versions of mature projects from a stable problem domain to increase the likelihood that we would encounter primarily maintenance activities in our studies.

@@

\section{Understanding Software Maintenance Activities} 
\label{sec:analysis}
% \todo[inline]{talk about this topic in one section}
% \todo[inline]{reference project name Nomen more}
%In order to test  our ROC values produced acceptable results, we now turn to
%\emph{applying} those techniques to understand software maintenance
%activities. 
%\todo[inline]{more on intro ... why should the reader care about
 % this. It needs to show how the stuff in the intro (applications) is
 % possible}
%%% XXX NEIL I don't get this.. VVVVV
%\todo[inline]{fix section headings}
As we mentioned in the introduction, a key issue in software maintenance is understanding \emph{why} a system has evolved the way it has~\cite{aranda09icse}. 
In this section we demonstrate the value of labelled topic extraction
in addressing this issue. 
% more why
Labelled topics address \emph{why}  because they show reasons why changes
occurred rather than \emph{how}. 
The \emph{how} of a change is the change itself, the purpose or
\emph{why} is what we are after.
We investigate the history of the two large-scale database systems that we studied. 
We use our technique to show the topic of development efforts over time in each project.
%We evaluated two research questions using the ground truth data that we annotated by hand:
We motivated our investigation with three research questions:
\begin{enumerate}
\item \emph{Do NFR frequencies change over time?} If a particular NFR
  was of more interest at one point in the life-cycle than another,
  this suggests that development activity shifted focus. For example,
  if a developer expected to see a recent focus on \emph{reliability},
  but instead \emph{usability} dominated, they might re-prioritize
  upcoming work items.

\item \emph{Do projects differ in their relative interest in NFRs?} A
  project manager, especially a systems-manager, would be interested
  in knowing whether a particular NFR, such as \emph{reliability}, was
  more important for one project than another. This could be to
  confirm the initial design goals, or to track the progress on that
  quarter's objectives.  The difference in NFR proportion is
  interesting because it implies a difference in focus between two projects.

\item \emph{Do different developers work on different NFRs?} For a given project, it is reasonable to think that developers are either assigned (in commercial organizations) or choose (in open-source organizations) to a particular NFR. For example, one developer might be more junior, and take responsibility for the low-impact reliability fixes. Another, more senior developer might assume responsibility for major improvements such as efficiency improvements.

\end{enumerate}

@@

%For MySQL (Figure \ref{fig:mysql-timeline}), we similarly validated our NFR patterns with external mailing list data. 
The release of MySQL we study  (Figure \ref{fig:mysql-timeline}) was the first to be licenced under the GPL. 
Version 3.23.31 (January, 2001) was the production release (non-beta), and we see a flurry of topics labelled with \emph{functionality} and \emph{maintainability}. 
After this point, this version enters the maintenance phase of its life-cycle. 
In May 2001, there is an increase in the number of topics labelled with \emph{portability}. 
This might be related to release 3.23.38, which focused on Windows compatibility. 
Similarly, in August, 2002, both \emph{functionality} and \emph{portability} are frequent, and mailing list data suggests this is related to the release of version 3.23.52, a general bug fix with a focus on security (a component of the \emph{functionality} NFR in the ISO9126 model). 
After this point, efforts shift to the newer releases (4.0, 4.1,
5.0). 
\todo{add PGSQL disucssion here. Schema additions? Lots of functionality?}

We now address our research questions:

@@

\subsection{Do NFR Frequencies Change over Time?}
%\noindent\textbf{1. Do NFR frequencies change over time?} --- 
%\subsubsection{Do label frequencies change over time?}
\todo{add more on PGSQL}
In both projects the frequencies generally decreased with age. 
However, there are variations within our NFR labels. In MySQL, \emph{usability} and \emph{efficiency} do not appear very often in topics. 
A proportionately smaller number of commits addressed these NFRs.
%As discussed in the previous section, 
Certain peaks in topic numbers coincide with a particular emphasis from the development team on issues such as new releases or bug fixes.
This suggests that maintenance activity is not necessarily strictly decreasing with time, but rather episodic and responsive to outside stimuli. 
In MaxDB, we can observe that \emph{Maintainability} topics became more prevalent as MaxDB matures. 
This is likely due to our analysis time-frame for MaxDB being shorter than the time-frame for the MySQL product.

@@

\subsection{Do different developers work on different NFRs?}
\todo{figure this out!}

>>>>

\begin{abstract}
% %\todo[inline]{Topics and LDA sooner, Topics/concept location is a known
Researchers have employed a variety of techniques to extract underlying
topics that relate to software development artifacts.  Typically, these
techniques use semi-unsupervised machine-learning algorithms to suggest
candidate word-lists.  However, word-lists are
difficult to 
interpret in the absence of meaningful summary labels.  Current topic modeling 
techniques assume manual labelling and do not use domain-specific knowledge
to improve, contextualize, or describe results for the developers.  
We propose a solution: \emph{automated labelled topic extraction}.
  Topics are extracted 
using  Latent Dirichlet Allocation (LDA) from commit-log comments recovered
from source control systems such as CVS and BitKeeper.  
These topics are given labels from a generalizable
cross-project taxonomy, consisting of non-functional
requirements. Our approach was evaluated with experiments and case studies
on two large-scale RDBMS projects: MySQL and MaxDB.  The case studies
show that labelled topic extraction can produce appropriate,
context-sensitive labels relevant to these projects, which provides fresh
insight into their evolving software development activities.

\keywords{ Software maintenance \and Repository mining \and Latent dirichlet allocation \and Topic models}

@@

To evaluate our approach, we sought candidate systems that were mature projects and had openly accessible source control repositories. 
We selected systems from the same application domain, to control for differences in functional, rather than non-functional, requirements. 
We selected MySQL 3.23 and MaxDB 7.500 as they were open-source, partially-commercial database systems. 
MaxDB started in the late 1970s as a research project, and was later acquired by SAP. 
As of version 7.500, released April 2007, the project has over $940,000$ lines of C source code\footnote{generated using David A. Wheeler's \emph{SLOCCount}, {http://dwheeler.com/sloccount}.}. 
The MySQL project started in 1994 and MySQL 3.23 was released in early 2001. 
MySQL contains $320,000$ lines of C and C++ source code.  
%Choosing an older version of these projects allowed us to focus on projects which have moved further into the maintenance phase of the software life-cycle.
We explicitly chose older versions of mature projects from a stable problem domain to increase the likelihood that we would encounter primarily maintenance activities in our studies.

@@

\section{Understanding Software Maintenance Activities} 
\label{sec:analysis}
% \todo[inline]{talk about this topic in one section}
% \todo[inline]{reference project name Nomen more}
%In order to test  our ROC values produced acceptable results, we now turn to
%\emph{applying} those techniques to understand software maintenance
%activities. 
%\todo[inline]{more on intro ... why should the reader care about
 % this. It needs to show how the stuff in the intro (applications) is
 % possible}
%%% XXX NEIL I don't get this.. VVVVV
%\todo[inline]{fix section headings}
As we mentioned in the introduction, a key issue in software maintenance is understanding \emph{why} a system has evolved the way it has~\cite{aranda09icse}. 
In this section we demonstrate the value of labelled topic extraction
in addressing this issue. 
% more why
Labelled topics address \emph{why}  because they show reasons why changes
occurred rather than \emph{how}. 
The \emph{how} of a change is the change itself, the purpose or
\emph{why} is what we are after.
We investigate the history of the two large-scale database systems that we studied. 
We use our technique to show the topic of development efforts over time in each project.
%We evaluated two research questions using the ground truth data that we annotated by hand:
We motivated our investigation with two research questions:
\begin{enumerate}
\item \emph{Do NFR frequencies change over time?} If a particular NFR
  was of more interest at one point in the life-cycle than another,
  this suggests that development activity shifted focus. For example,
  if a developer expected to see a recent focus on \emph{reliability},
  but instead \emph{usability} dominated, they might re-prioritize
  upcoming work items.

\item \emph{Do projects differ in their relative interest in NFRs?} A
  project manager, especially a systems-manager, would be interested
  in knowing whether a particular NFR, such as \emph{reliability}, was
  more important for one project than another. This could be to
  confirm the initial design goals, or to track the progress on that
  quarter's objectives.  The difference in NFR proportion is
  interesting because it implies a difference in focus between two projects.
\end{enumerate}

@@

%For MySQL (Figure \ref{fig:mysql-timeline}), we similarly validated our NFR patterns with external mailing list data. 
The release of MySQL we study  (Figure \ref{fig:mysql-timeline}) was the first to be licenced under the GPL. 
Version 3.23.31 (January, 2001) was the production release (non-beta), and we see a flurry of topics labelled with \emph{functionality} and \emph{maintainability}. 
After this point, this version enters the maintenance phase of its life-cycle. 
In May 2001, there is an increase in the number of topics labelled with \emph{portability}. 
This might be related to release 3.23.38, which focused on Windows compatibility. 
Similarly, in August, 2002, both \emph{functionality} and \emph{portability} are frequent, and mailing list data suggests this is related to the release of version 3.23.52, a general bug fix with a focus on security (a component of the \emph{functionality} NFR in the ISO9126 model). 
After this point, efforts shift to the newer releases (4.0, 4.1,
5.0). We now address our research questions:

@@

\subsection{Do NFR Frequencies Change over Time?}
%\noindent\textbf{1. Do NFR frequencies change over time?} --- 
%\subsubsection{Do label frequencies change over time?}
In both projects the frequencies generally decreased with age. 
However, there are variations within our NFR labels. In MySQL, \emph{usability} and \emph{efficiency} do not appear very often in topics. 
A proportionately smaller number of commits addressed these NFRs.
%As discussed in the previous section, 
Certain peaks in topic numbers coincide with a particular emphasis from the development team on issues such as new releases or bug fixes.
This suggests that maintenance activity is not necessarily strictly decreasing with time, but rather episodic and responsive to outside stimuli. 
In MaxDB, we can observe that \emph{Maintainability} topics became more prevalent as MaxDB matures. 
This is likely due to our analysis time-frame for MaxDB being shorter than the time-frame for the MySQL product.

@@
